---
title: "Data Visualization"
output: html_document
---

<br/>
<br/>

### Data Visualization with `ggplot2`

`ggplot2` is a package for plotting data. The "gg" stands for "grammar of graphics". To use `ggplot2`, first you must install and load the package.

``` {r}
install.packages("ggplot2", repos = "http://cran.us.r-project.org/")
library(ggplot2)
```

<br/>
<br/>

The code for creating a plot using ggplot is more complicated than using `boxplot()` or `hist()`. Here's an example that we'll walk through step by step:

``` {r}
# We're going to use the 'describe' function from the psych package, so first load it:
library(psych)

# Let's create a random data set with 50 observations of reaction time in two conditions.
mydata <- data.frame(id = seq(1:50),  # create a column named "id" with the numbers (sequence) 1 through 50
                     condition = rep(c("A", "B"), 25),  # create a "condition" column with A and B repeating
                     rt = rnorm(n = 50, mean = 100, sd = 15))  # create a third column named "rt" with randomly generated data

# Now, get the mean and standard error for each condition
mydata.summary <- describeBy(mydata, "condition")
mydata.summary
mydata.plot <- data.frame(condition = factor(c("A", "B")),
                          meanRT = c(mydata.summary[[1]]$mean[3], mydata.summary[[2]]$mean[3]),
                          se = c(mydata.summary[[1]]$se[3], mydata.summary[[2]]$se[3]))

ggplot(mydata.plot, aes(x = condition, y = meanRT)) +
  geom_bar(stat= "identity", width = .5) +
  geom_errorbar(aes(ymin = meanRT - se, ymax = meanRT + se), width = .25)
```

<br/>
<br/>

Let's go through the code step by step.

<br/>

First, we load the `psych` package and create our data.frame. 

```
# We're going to use the 'describe' function from the psych package, so first load it:
library(psych)

# Let's create a random data set with 50 observations of reaction time in two conditions.
mydata <- data.frame(id = seq(1:50),  # create a column named "id" with the numbers (sequence) 1 through 50
                     condition = rep(c("A", "B"), 25),  # create a "condition" column with A and B repeating
                     rt = rnorm(n = 50, mean = 100, sd = 15))  # create a third column named "rt" with randomly generated data
```

Then, we must calculate the mean reaction time and standard error of reaction time for each condition.

``` {r}
mydata.summary <- describeBy(mydata, "condition")
mydata.summary
```

The `describeBy` function calculates a bunch of summary statistics for our data.frame, and does it separately for each condition. The result is two tables, one for each condition.

<br/>

We need to store the mean and standard error for the reaction time data in each condition. That's what the next 3 lines of code do:

```
mydata.plot <- data.frame(
```

We're going to store what we need in a data.frame called "mydata.plot"

```
condition = factor(c("A", "B")),
```

The first column lists the two conditions, A and B. (Our data.frame will have 2 rows.)

```
meanRT = c(mydata.summary[[1]]$mean[3], mydata.summary[[2]]$mean[3]),
```

We're combining two things and storing them in a column named "meanRT". What do all of those brackets and numbers mean?

<br/>

The mydata.summary object, which was created by the describeBy function, is a special type of class called a List. Lists are like vectors, only instead of each element storing a single value, you can store many objects in a list - for example, more than one data.frame.

<br/>

Just like with vectors, we can index lists using brackets:

``` {r}
mydata.summary[1]
class(mydata.summary[1])
```

Using a single set of brackets will let us select the first item in the list, but it's still of the class 'list'. It would be great if we could treat that table as a data.frame, so that we can work with it using the other things we know. That's exactly what the double set of brackets does:

``` {r}
mydata.summary[[1]]
class(mydata.summary[[1]])
```

You'll see that the resulting object has three classes - "psych", "describe", and most useful for us, "data.frame". Now, we can use the $ symbol to select a column by name.

```{r}
mydata.summary[[1]]$mean
```

This selects the column of means for condition A. We're interested in only the mean reaction time, so we can again use brackets to select the third element:

```{r}
mydata.summary[[1]]$mean[3]
```

This code says to select the third element of the column named "mean" of the content of the first item in the list "mydata.summary".

<br/>

Now, we can change the 1 to a 2 to select the mean reaction time for condition B, and combine them using `c()`.

``` {r}
c(mydata.summary[[1]]$mean[3], mydata.summary[[2]]$mean[3])
```

We store the result in a column of our new data.frame named "meanRT". We can do the exact same thing to select the standard error (se) of reaction time for each condition:

```
se = c(mydata.summary[[1]]$se[3], mydata.summary[[2]]$se[3]))
```

We end up with a data.frame that has two rows and three columns:

``` {r}
mydata.plot
```

<br/>

Now let's break down the ggplot code:

```
ggplot(mydata.plot, aes(x = condition, y = meanRT)) +
```

The `ggplot()` function creates an object of the class "ggplot". Here, we give it two arguments. The first tells it the name of our data.frame (mydata.plot), and the second tells it how to arrange the data on the x and y axes. We want two columns representing our conditions arranged along the x-axis, and the height of the bars to correspond to the mean reaction time in each condition. We use `aes()` to indicate this (it stands for _aesthetic mapping_).

<br/>

We've created a ggplot object with condition on the x-axis and mean reaction time in the y-axis, but we haven't specified what type of plot we want - a bar plot? Line plot? Something else? So, we add more to the ggplot object with the plus sign.

```
geom_bar(stat= "identity", width = .5) +
```

`geom_bar()` indicates that we want a bar plot. We provide two arguments: `stat = "identity"` indicates that the height of the bars should be exactly what's in the meanRT column, and `width = .5` sets the width of the bars to 50% of the default width.
  
```
geom_errorbar(aes(ymin = meanRT - se, ymax = meanRT + se), width = .25)
```

And finally, we add another line to add errorbars to the plot using the `geom_errorbar()` function. Importantly, we need to specify how wide the bars should be using the `aes()` argument. Inside `aes()`, we set the minimum y-value for the bars to be the mean minus the standard error (meanRT - se), and the maximum y-value to be the mean plus the standard error (meanRT + se). We can also set the width using another argument.

<br/>
<br/>

`ggplot2` can make many other complex visualizations. You can read more about how to use it at http://docs.ggplot2.org/current/
