---
title: "Indexing and Summarizing Data - Homework "
output: html_document
---

Main Problems
-------------

1. Create a vector named "odd" with all odd numbers from 1 - 13.

``` {r}
odd <- c(1, 3, 5, 7, 9, 11, 13)
```

<br/>

2. Select the third element and raise it to the 5th power.

``` {r}
odd[3] ^ 5
```

<br/>

3. Create another vector named "even" with all even numbers from 2 - 14.

``` {r}
even <- c(2, 4, 6, 8, 10, 12, 14)
```

<br/>

4. Create a data.frame named "numbers" with two columns: one named "apple" that contains your vector of odd numbers, and one named "orange" that contains your vector of even numbers.

``` {r}
numbers <- data.frame(apple = odd, orange = even)
```

<br/>

5. Select all rows where "orange" is greater than 7.

``` {r}
numbers[numbers$orange > 7, ]
```

<br/>

6. Calculate the mean of the full "apple" column.

``` {r}
mean(numbers$apple)
```

7. Without using the `$`, create a third column called `total` that is apples plus oranges

```{r}

```

Bonus
-----

### The `seq` function

Based on the output of the code below, how would you generate 6 evenly spaced numbers, from 1 to 10.

```{r}
seq(from=1, to=5, by=2)
seq(from=1, to=5, length.out=3)
seq(from=1, to=5, length.out=2)
# try it here
```

How would you do it using the other option (either the `by` or `length.out` argument)?

```{r}



```

### Recycling

The following might exhibit some surprising behavior, due to something called recycling..

```{r, error=TRUE}
x <- 1:4
x[c(TRUE, FALSE, FALSE, FALSE)]
x[c(TRUE, TRUE, TRUE, TRUE)]
x[TRUE]
x[c(TRUE, FALSE)]
x[c(TRUE, FALSE, FALSE)]
```

The data.frame function might offer some hints..

```{r}
data.frame(x = x, indx = c(TRUE, FALSE))
```

How can you get every 3 letters in the vector below (that is, `A, D, ..., Y`) using the fewest number of TRUE or FALSE possible. 
Note that the `LETTERS` variable is built in to R, and is a vector with A to Z.

```{r}
LETTERS[1:10]
# modify line below
LETTERS[c(TRUE)]

```

### The `paste` and `rep` functions

**One interesting function that uses the concept of recycling is `paste`.**

```{r}
paste(LETTERS[1:4], 1:4)

paste(LETTERS[1:4], 1:3)

paste('the letter', LETTERS[1:4])

paste('this letter is', LETTERS[1:4], sep=": ")
```

**Another which is often useful in conjunction with `paste` is `rep`.**

```{r}
ltrs <- LETTERS[1:4]
rep(ltrs, times=2)
rep(ltrs, each=2)
rep(ltrs, length.out=10)
```

How could you make the string counting up from 1 to 5 by .5 (that is, `1, 1.5, 2, ..., 5`), but label each number with it's whole part. For example, the first three entries should be: `"one: 1", "one: 1.5", "two: 2"`?

```{r}

```
