---
title: "t-test and anova homework"
output: html_document
---

Main Problems
-------------

1. R has a function called `options` that let's you configure different settings.
   One particularly useful setting is digits. 
   As you do this homework, modify the number of digits and see what happens.

```{r}
options(digits=7)
```

2. Read in the working memory training data `training_long.csv` in the data folder.

```{r}
dat <- read.csv('../data/training_long.csv')
```

3. Make a boxplot of the pre-test data.

```{r}
pre <- subset(dat, time == "pre")
boxplot(wm_score ~ cond, data=pre)
```


3. Report the relevant descriptives, anova, and Tukey's HSD for the pre-test data.

```{r}
library(psych)

# descriptives
with(pre, describeBy(wm_score, cond, mat=TRUE))

# anova
fit <- aov(wm_score ~ cond, data=pre)
anova(fit)

# Tukey HSD
tuk <- TukeyHSD(fit)
tuk
```

4. Use an inline chunk to report the p-value from the wm vs combo t-test in Tukey's HSD. The p-value is `r tuk$cond['wm-combo', 'p adj']`.

Reshaping
---------

Another interesting use of formulas is in the `reshape2` package. 
This package focuses on how you can represent your data. 
For example, the training data has a separate row for the post- and pre-test scores, so each subject's data takes two rows.

```{r}
dat <- read.csv('../data/training_long.csv')
subset(dat, subject == 1)
```

However, we can "cast" it so that rather than being separate rows, `pre` and `post` have their own columns on a single row. 
Essentially, we we're saying "make each `subject` their own line, but make `time` the columns". 
In this sense, the formula is `rows ~ columns`

```{r}
library(reshape2)
dcast(dat, subject ~ time, value.var="wm_score")     # the d stands for data.frame!
```

1. Below, I set up a data set that already exists in R called `ChickWeight`. Which of the two variables in this data could be cast into columns? Which should be the `value.var`? Try it below.

```{r}
# ChickWeight is a built in data set
ChickWeight$Diet = NULL
head(ChickWeight)

```

2. Below I've read in and modified `training_long.csv`. In this simulation, each subject completed all 3 training conditions. In this sense, both condition **and** time can be cast on the columns. Given what you've learned about formulas, try to modify the formula from the example to do this below.

```{r}
train_long <- read.csv('../data/training_long.csv')
train_long$subject <- 1:10

# cast here
train_wide <- dcast(train_long, subject ~ cond + time, value.var = "wm_score")
```

3. Now that each subject has a row, you can use the `plot` function to make scatterplots. Use indexing to remove the subject column, then try `plot(YOURDATANAME)`.

```{r}
plot(subset(train_wide, select=-subject))
# or
plot(train_wide[-ncol(train_wide)])
```

4. Plot uses formulas, too. Based on what you've learned about formulas, replace `LEFT` and `RIGHT` below to make a scatter plot.

(remember, `a ~ b` can be read "a by b")

```{r}
x <- rnorm(8)
wide <- data.frame(subject = 1:8,
                     x = x,
                     y = x + rnorm(8))
long <- melt(wide, id.vars='subject', variable.name='group')
long
plot(value ~ group, data=long)
```

References
----------

[simple reshape2 tutorial](http://seananderson.ca/2013/10/19/reshape.html)
[original reshape2 article](http://www.jstatsoft.org/v21/i12)
[reshape2 home page]