---
title: "Common Errors and Debugging"
output: html_document
---

```{r setup, include=FALSE}
# this chunk lets us include errors in our code
knitr::opts_chunk$set(error=TRUE)
```

Variable does not exist
-----------------------
```{r}
a + 1
```

Calling a function without the necessary arguments
--------------------------------------------------

For example, using the `with` function requires something like `with(data, expr)`.
Here, `data` might be the data.frame you're using, and `expr` will be an expression that uses its columns.

```{r}
dat <- data.frame(a=1:10)
with(dat, a + 1)    # good
with()              #bad
```


You tried to refer to your data, but used the wrong variable name
-----------------------------------------------------------------

This is a very common error that can produce cryptic errors.
For example, you might have a variable called `dat`, but you accidentally try to refer to it as `df` or `data`, which are actually R functions!

```{r}
dat <- data.frame(a=1:6)
colMeans(dat)

colMeans(df)
colMeans(data)
```

Because `df` and `data` are functions, you can get a lot of different errors, depending on what you try to do with them.

```{r}
# try to subset
subset(df, a < 5)
# try to access column
df$a
# try to make boxplot
boxplot(df)
```

On the topic of subset, you might not get an error, but it may behave unexpectedly.


```{r}
b <- 3

subset(dat, b < 5)
```

Notice that `dat` doesn't have a `b` column, so R used the variable `b` that we created outside of the data.frame. 
This can create problems if you mispell the name of a column in a data.frame, but your mispelling is the name of an existing varible.

You try to use recycling when creating a data.frame, but the vector lengths don't divide evenly
----------------------------------------------------------------------------------------

R will recycle values when creating a data.frame. This means that you can do something like..

```{r}
# works since 4 / 2 is 2
data.frame(a=1:4, b=1:2)

# doesn't work since 4 / 3 is not a whole number
data.frame(a=1:4, b=1:3)
```

You try to add things that aren't numbers
--------------------------------------

```{r}
1 + "a"
"a" + "a"

fact <- factor(c('a','a','b'))
fact + fact
```

Note that the last expression, adding two factors only returned a warning.

You data has NA or NaN values
-----------------------------

These values can be produced when performing an operation that isn't defined, like `sqrt(-1)`. Many functions have an `na.rm` argument that will ignore `NA` and `NaN`

```{r}
sqrt(-1)

NA + 1       # NA

sum(sqrt(-1:10))

# use the na.rm argument
sum(sqrt(-1:10), na.rm=TRUE)
```

