---
title: "Phonological Similarity and Memory Data"
output: html_document
params:
  tasks: 
    label: 'Task'
    value: 'Rspan.long'
    input: 'select'
    choices: [R.scramb, Rspan.long, Rspan.short, Rspan.names.short, Rspan.names.long]
    multiple: TRUE
  dv: 'ACC.ser'
---

This data was collected as part of a set of experiments on memory. 
In the experiments, participants switched between memorizing words, and reading irrelevent sentences.
After being shown a certain number of to-be-remembered (**TBR**) words to memorize, participants were asked to recall as many of those words as possible. For example, if a trial had 2 words to memorize, participants would..

see TBR word, read sentence, see new TBR word, read new sentence, recall words.

Researchers were interested in the effect of **task** and **trialtype** on recall performance.

The Data
--------

The data has 8 columns..

* **task:** the experimental condition
* **Subject:** id for each unique subject within a task
* **trialnum:** 4 to 27 for each of the 27 trials participants completed (the first 3 were practice)
* **trialtype:** the type of words participants needed to memorize on that trial
* **TBR:** the to-be-rememberd word
* **inpos:** the order of the TBR word (e.g. 0 was the first word shown)
* **ACC.ser:** did they recall the right word in the right position?
* **ACC.item:** did they recall the right word?

```{r}
dat <- read.csv('../data/scored_recall.csv')
subset(dat, Subject == 8 & trialnum == 5)
```

Using rmarkdown params
----------------------

Here, we subset the data to only the tasks specified in params.
The if condition let's us ignore this portion if we aren't knitting our document.

```{r}
# exists checks if the variable has been defined
if (exists('params')) {
  dat <- subset(dat, task %in% params$tasks)
} else {
  # could create a params list here, e.g.
  # params = list(tasks = c(), dv = "")
}
```


Pre-processing
--------------

Note that in on the bottom row, `inpos` is `NA`. This means that the participant recalled more words than were shown to them.
The first thing we'll want to do is remove those `NA` rows, so that we're only considering valid recalls.

```{r}
dat <- subset(dat, !is.na(inpos))
```

Next, we'll average accuracy within each 
```{r}
# Can change the line below to configure choosing which variable to use
dat$dv = dat$ACC.ser
#
dat_agg <- aggregate(dv ~ task + Subject + trialtype, data=dat, mean)
head(dat_agg)
```

It's a little bit annoying that it isn't ordered by task and Subject anymore.
We can fix that using the `order` function.
```{r}
row_ord <- with(dat_agg, order(task, Subject, trialtype))
dat_ord <- dat_agg[row_ord,]
```

Finally, let's reduce our scores down to a single mean and standard deviation for each condition and trialtype.

```{r}
se <- function(x) { 
  sqrt(var(x) / length(x)) 
}
means <- aggregate(dv ~ task + trialtype, dat_ord, mean)
ses <- aggregate(dv ~ task + trialtype, dat_ord, se)

meanse <- means
meanse$se = ses$dv
head(meanse)
```


Plotting
--------

```{r}
library(ggplot2)

# dodge is useful for moving points to horizontally so they don't overlap
dodge = position_dodge(width=.1)
# plot points with errorbars
# we give all of the aesthetics we want as defaults to the ggplot function
# and all additional aesthetics to specific geoms
ggplot(meanse, aes(task, dv, color=trialtype)) + 
  geom_point(position=dodge) +
  geom_errorbar(aes(ymin=dv-se, ymax=dv+se), width=.1, position=dodge)

```

