---
title: "homework for 2c — advanced R"
output: html_document
---

1. Write an if statement that checks if `a` is a character, and if it is, changes it to an integer (using as.integer).

```{r}

a <- c('5', '6', '10')

if (is.character(a)) {
  a <- as.integer(a)
}

```

2. Write a for loop that adds each number in `b` to the number before it.

```{r}

b <- c(1:3)

for (ii in 2:length(b)) {
  b[ii] = b[ii] + b[ii-1]
}
# the resulting array should be c(1, 3, 5)
```

3. Write a function that takes a vector `x`, and adds 10 to it (be sure to return `x`). Call it using `c`.

```{r}
# uncomment line below and define function
 f <- function(x) {
   x + 10
 }

# call function below
c <- 1:3
f(c)
```

4. Let's connect the previous three ideas together. Write a function, `adder`, that takes an argument, `x`. This function should..
  * check if `x` is an integer, and if it isn't then make it one.
  * use a for loop to add the each number to the number before it
  * return the array with the new numbers in it
  
```{r}
function(x){
  # make x an integer
  if (is.character(x)) x = as.integer(x)
  # loop over x
  for (ii in 2:length(x)) x[ii] = x[ii] + x[ii-1]
  # return x
  x
}
```

