---
title: "Basics of R"
output: html_document
---

<br/>
<br/>

### Arithmetic operators

You can do mathematical operations in R.

``` {r}
2 + 2
10 - 2.34
5 * 12
99 / 11
2^4
(100 - 13) * 3^2
```

<br/>

Spacing is not very important.

``` {r}
2    +  2
3*4
```

<br/>
<br/>

### Objects and the Assignment Operator <-

In R, you can store data, statistics, and other information in __objects__. Objects are represented by a variable name, and are created by using the assignment operator <-

```{r}
anumber <- 2 + 2  # We're storing the result of 2+2 in the object "anumber"
anumber  # This will print the object.
# Note: The # symbol begins a comment, and everything following a # will not affect your code.
```
You can also use = instead of <- ...
``` {r}
anumber = 9.123
anumber
```
but it's better to use <- to differentiate between assigning something to an object and other uses of =, which we'll get to later. You can think of the leftwards arrow as storing data _in_ the object, rather than just setting a variable equal to some value.

<br/>
<br/>

### Object Classes with class( )

Every object in R has a __class__, or a type that describes the content and organization of an object.

``` {r}
class(anumber)
```

Our object 'anumber' is of the class __numeric__. There are a few other classes of object in R:

``` {r}
alogical <- TRUE  # This class is called "logical", and takes the value TRUE or FALSE
class(alogical)
acharacter <- "data is"  # This class is called "character", and holds strings of text
class(acharacter)
afactor <- factor("A")  # This is a class called "factor". It's like character, but when elements with the same text are grouped
class(afactor)
```

... plus many other more complex classes, some of which we'll cover later.

<br/>
<br/>

### Changing object classes

Sometimes you need to change the class of an object. You can _coerce_ some object classes by using as.numeric( ), as.logical( ), as.character( ), and as.integer( ).

``` {r}
a <- "1"
class(a)
as.integer(a)
class(as.integer(a))

b <- 14
class(b)
as.character(b)
class(as.character(b))
as.factor(b)
class(as.factor(b))

c <- 1
d <- 0
class(c)
class(d)
as.logical(c)
as.logical(d)
class(as.logical(c))
class(as.logical(d))
```

<br/>
<br/>

### Combining things with c( )

We use c( ) to _combine_ or _concatenate_  __elements__ (numbers, integers, logical values, or character strings):

``` {r}
somenumbers <- c(5.234, 3, 10.1, 900)
somelogical <- c(TRUE, TRUE, FALSE, TRUE, FALSE)
somecharacter <- c("Princeton", "Yale", "Harvard", "Penn")
somefactor <- factor(c("A", "B", "C", "B", "B", "A"))
somenumbers
somelogical
somecharacter
somefactor
```

c( ) is a __function__ that combines various arguments or __elements__ to form a __vector__, a one-dimensional series of values. 

<br/>

A function is like a command that takes several __arguments__ and returns a result. In R, you call functions with a function name followed by a set of parentheses. Inside of the parentheses go what are called arguments, which indicate various details about how the function should operate. In the c( ) function, the arguments of inside the parentheses after the c are the things to be combined.

<br/>

We've used a few other functions so far. class( ) is a function that returns the class of the object in parentheses. as.integer( ) and the other coercion functions change the class of the object in parentheses. factor( ) encodes its argument as a factor.

<br/>

There are many __base__ functions in R, i.e. functions that are built-in and ready to use. Sometimes, you might need to do something that isn't easily done with the base functions. In R, you can install __packages__, which include additional functions written by other R users. You can install packages using the install.packages( ) function, and load them using library( ).

<br/>

This code installs and loads the "psych" package, which provides some functions that are useful for doing statistical analyses of experimental data.

``` {r}
install.packages("psych")
library("psych")
```

<br/>
<br/>

#### Tip: Help with functions

You can access the help page for a function by using the ? mark. This will show you information about what the function does and which arguments you can specify. Try typing ?c into the console. ? will try to access the help page for a particular command, while ?? will search the help pages.

<br/>
<br/>

### Recap: Objects, Classes, Vectors, & Functions

To recap, information we work with in R is stored in __objects__. Each object has one or more __class__ (or type). Four types of __vectors__ (one dimensional objects) are:

* __Numeric__: numbers with decimal values
* __Integers__: numbers without decimal values
* __Logical__: True or False
* __Character__: strings of text
* __Factor__: strings of text that indicate groups, conditions, etc.

__Functions__ are commands that execute code, and contain __arguments__ that detail how to do so.

<br/>
<br/>

### data.frames - Another kind of object

A data.frame is another kind of object, where data are stored in a matrix or table-like format. Let's try creating a simple data frame with one row and three columns for each participant in a research study: "id" contains a unique number to identify each participant, "condition" indicates whether each participant was assigned to condition A or condition B, and "rt" is the reaction time in a task.

You'll see some new functions in the code below. 

``` {r}
# We use data.frame() to create a new data.frame from scratch.
mydata <- data.frame(id = seq(1:50),  # create a column named "id" with the numbers (sequence) 1 through 50
                     condition = rep(c("A", "B"), 25),  # create a "condition" column with A and B repeating
                     rt = rnorm(n = 50, mean = 100, sd = 15))  # create a third column named "rt" with randomly generated data
class(mydata)
```

Now we've created a data.frame with 50 rows and 3 columns. You can examine the first few rows by using the head( ) function:

``` {r}
head(mydata)
```

You can also view the full data.frame itself by entering "mydata", or you can view it in a table by entering View(mydata) or clicking on the mydata object in the Environment tab in RStudio.

<br/>
<br/>

### Reading data from a file

Often, the data you want to analyze is located in a separate file. To work with it in R, you need to _read_ in the data.

<br/>

There are a few different functions for reading data into R, depending on the type of data file.

``` {r}
#mydata <- read.csv() # read.csv() is used for Comma Separated Value files
#mydata <- read.delim() # read.delim() is used when your data are separated/delimited by the TAB character
#mydata <- read.table()
```
