---
title: "Basics of R - Homework "
output: html_document
---

<br/>
<br/>

1. Create a vector named "zip" with 5 elements, each one a digit of Princeton's zip code (08544).

``` {r}

# Replace this comment with your code


```

<br/>

2.  Check the class of "zip".

``` {r}

```

<br/>

3. Install the "effsize" package (but set the code chunk to not **eval**uate).

``` {r}

```

4. Read in `finch-beaks.csv` from the data folder, assign it to the variable name `dat`. Then, use the `summary` function on `dat`.

```{r}

```

**bonus**

The finch beak data were collected on an island in the Gallapagos by two researchers from Princeton, who measured the beak size of finches. The data contains two columns.

* Year -- when the measurement was taken
* Depth -- the size of the beak

If you named your data correctly above, you should be able to look at, for example, the `Year` column by using `dat$Year`. 
Try out some of the following functions on one of these columns: `boxplot`, `histogram`, `range`.

```{r}


```

Now, try reading in `finch-beaks.tsv` in the data folder

```{r}

```

